---
title: "distribution fitting"
author: "Zach Peagler"
date: "10/12/2024"
output: html_document
---

## Load Packages
```{r packages, include=FALSE}
require(dplyr); require(ggplot2); require(MASS);
```

## Graphical Setup
```{r gsetup}
a_palette <- "lipari"
## fonts
font_add_google("Open Sans", family = "open")
font_add_google("Montserrat", family = "mont")
showtext_auto()
```

##### REMOVE BEFORE PUBLICATION #####

## Test Data
```{r test_data}

dat <- read.csv("C:/Github/App-A-Day/02_distribution_fitter/eggplant_data.csv")

```

#######################################

```{r distributions}

# (-inf:inf)

# (0:inf)
cont_distributions <- function() {
  return(c("normal", "lognormal", "gamma", "exponential", "cauchy", "t", "weibull", "logistic"))
}

# (0-1)
## beta, 

#  if ("beta" %in% distributions) {
#    x_beta <- fitdist(x, "beta")
#  }

# geometric / number of fails in a sequence of Bernoulli trials before success occurs

#  if ("geometric" %in% distributions) {
#    x_geom <- fitdistr(x, "geometric")
#  }

```

## Continuous PDF
### Description
This is used to test the distribution of non-negative continuous variables. It returns a dataframe (table) with the a  and the predicted distribution based on.
### Usage

> cont_pdf(x, seq_length, distributions, ...)

### Arguments
x <- the variable that is being fitted against distributions. must be continuous and non-negative.
seq_length <- the length of the sequence used to generate distributions. results in the number of rows in the final output.
distributions <- a list of distributions to test against. options include "normal", "lognormal", "Gamma", "exponential", "beta", "cauchy", "t", "weibull", "logistic", and "geometric". calling "all" will return all possible distributions.

### Details

Uses sequence length via seq() on the min and max of x over seq_length + 1. 

Then computes the kernel density of x with density().

Create a new dataframe out of the x sequence and computed kernel density.

Check distributions against each supported distribution. If requested (and capable of doing so; continuous non-negative numbers required.) it fits the distribution of x against the requested distribution and adds it to the returned dataframe.

```{r cont_pdf}

cont_pdf <- function(x, seq_length, distributions){
  # get a sequence from the minimum to maximum of x with length
  #equal to seq_length + 1
  x_seq <- seq(min(x), max(x), length.out = seq_length+1)
  # create real density for x
  x_pdf <- density(x, n=seq_length+1)
  # initialize df of x and the real density
  pdf_df <- as.data.frame(x_seq) %>%
    mutate(dens = x_pdf$y)
  if ("all" %in% distributions) {
    distributions <- c("normal", "lognormal", "gamma", "exponential",
                       "cauchy", "t", "weibull", "logistic")
  }
  
  if ("normal" %in% distributions) {
    x_n <- fitdistr(x, "normal")
    x_pdf_n <- dnorm(x_seq, mean=x_n$estimate[1],
                   sd = x_n$estimate[2])
    pdf_df <- pdf_df %>% mutate(pdf_normal = x_pdf_n)
  }
  if ("lognormal" %in% distributions) {
      x_ln <- fitdistr(x, "lognormal")
      x_pdf_ln <- dlnorm(x_seq, meanlog=x_ln$estimate[1],
                   sdlog = x_ln$estimate[2])
      pdf_df <- pdf_df %>% mutate(pdf_lognormal = x_pdf_ln)
  }
  if ("gamma" %in% distributions) {
    x_g <- fitdistr(x, "gamma")
    x_pdf_g <- dgamma(x_seq, shape=x_g$estimate[1],
                    rate=x_g$estimate[2])
    pdf_df <- pdf_df %>% mutate(pdf_gamma = x_pdf_g)
  }
  if ("exponential" %in% distributions) {
    x_exp <- fitdistr(x, "exponential")
    x_pdf_exp <- dexp(x_seq, rate = x_exp$estimate)
    pdf_df <- pdf_df %>% mutate(pdf_exponential = x_pdf_exp)
  }
  if ("cauchy" %in% distributions) {
    x_cau <- fitdistr(x, "cauchy")
    x_pdf_cau <- dcauchy(x_seq, location=x_cau$estimate[1],
                         scale = x_cau$estimate[2])
    pdf_df <- pdf_df %>% mutate(pdf_cauchy = x_pdf_cau)
  }
  if ("t" %in% distributions) {
    x_t <- fitdistr(x, "t")
    x_pdf_t <- dt(x_seq, df = x_t$estimate[3])
    pdf_df <- pdf_df %>% mutate(pdf_t = x_pdf_t)
  }
  if ("weibull" %in% distributions) {
    x_wei <- fitdistr(x, "weibull")
    x_pdf_wei <- dweibull(x_seq, shape = x_wei$estimate[1],
                    scale = x_wei$estimate[2])
    pdf_df <- pdf_df %>% mutate(pdf_weibull = x_pdf_wei)
  }
  if ("logistic" %in% distributions) {
    x_logis <- fitdistr(x, "logistic")
    x_pdf_logis <- dlogis(x_seq, x_logis$estimate[1],
                          x_logis$estimate[2])
    pdf_df <- pdf_df %>% mutate(pdf_logistic = x_pdf_logis)
  }

  return(pdf_df)
}

```

## ggPDF
### Description
A function to return a ggplot showing the probability distribution function PDF of a variable.
### Usage
### Arguments
### Details
```{r ggpdf}

## Example plot
ggplot(cont_pdf(dat$Fv.Fm, 100, "all")) +
    geom_path(aes(x=x_seq, y=dens, color = "Actual Density"), size=3)+
    geom_path(aes(x=x_seq, y=pdf_normal, color="Normal"), size=2)+
    geom_path(aes(x=x_seq, y=pdf_lognormal, color="Lognormal"),size=2)+
    geom_path(aes(x=x_seq, y=pdf_gamma, color="Gamma"), size=2)+
    geom_path(aes(x=x_seq, y=pdf_exponential, color="Exponential"),size=2)+
    geom_path(aes(x=x_seq, y=pdf_cauchy, color="Cauchy"), size=2)+
    geom_path(aes(x=x_seq, y=pdf_t, color="t"),size=2)+
    geom_path(aes(x=x_seq, y=pdf_weibull, color="Weibull"), size=2)+
    geom_path(aes(x=x_seq, y=pdf_logistic, color="Loglogistic"), size=2)+
    theme_bw()+
    scale_color_scico_d(begin=0, end=0.8, palette = a_palette)+
    scale_shape_discrete()+
    guides(color = guide_legend(title="Distribution"))+
  labs(title = "Fv/Fm Probability Distribution Function (PDF) Plot")+
    ylab("PDF")+
    xlab("Fv/Fm")+
    theme(
      text = element_text(size=24, family="mont"),
      legend.position="right",
      legend.position.inside = c(.85,.6),
      axis.title = element_text(size=24, family = "mont", face= "bold"),
      title = element_text(size=24, family = "open", face= "bold")
    )

```

## Continuous CDF function
### Description
### Usage
### Arguments
### Details
```{r cont_cdf}

cont_cdf <- function(x, seq_length, distributions){
  # get a sequence from the minimum to maximum of x with length
  #equal to seq_length + 1
  x_seq <- seq(min(x), max(x), length.out = seq_length+1)
  # create real cumulative density for x
  x_cdf <- ecdf(x)(x_seq)
  # initialize df of x and the cumulative density
  cdf_df <- as.data.frame(x_seq) %>%
    mutate(dens = x_cdf)
  if ("all" %in% distributions) {
    distributions <- c("normal",
                       "lognormal",
                       "gamma",
                       "exponential",
                       "cauchy",
                       "t",
                       "weibull",
                       "logistic")
  }
  
  if ("normal" %in% distributions) {
    x_n <- fitdistr(x, "normal")
    x_cdf_n <- pnorm(x_seq, mean=x_n$estimate[1],
                   sd = x_n$estimate[2])
    cdf_df <- cdf_df %>% mutate(cdf_normal = x_cdf_n)
  }
  if ("lognormal" %in% distributions) {
      x_ln <- fitdistr(x, "lognormal")
      x_cdf_ln <- plnorm(x_seq, meanlog=x_ln$estimate[1],
                   sdlog = x_ln$estimate[2])
      cdf_df <- cdf_df %>% mutate(cdf_lognormal = x_cdf_ln)
  }
  if ("gamma" %in% distributions) {
    x_g <- fitdistr(x, "gamma")
    x_cdf_g <- pgamma(x_seq, shape=x_g$estimate[1],
                    rate=x_g$estimate[2])
    cdf_df <- cdf_df %>% mutate(cdf_gamma = x_cdf_g)
  }
  if ("exponential" %in% distributions) {
    x_exp <- fitdistr(x, "exponential")
    x_cdf_exp <- pexp(x_seq, rate = x_exp$estimate)
    cdf_df <- cdf_df %>% mutate(cdf_exponential = x_cdf_exp)
  }
  if ("cauchy" %in% distributions) {
    x_cau <- fitdistr(x, "cauchy")
    x_cdf_cau <- pcauchy(x_seq, 
                         location = x_cau$estimate[1],
                         scale = x_cau$estimate[2])
    cdf_df <- cdf_df %>% mutate(cdf_cauchy = x_cdf_cau)
  }
  if ("t" %in% distributions) {
    x_t <- fitdistr(x, "t")
    x_cdf_t <- pt(x_seq, df = x_t$estimate[3])
    cdf_df <- cdf_df %>% mutate(cdf_t = x_cdf_t)
  }
  if ("weibull" %in% distributions) {
    x_wei <- fitdistr(x, "weibull")
    x_cdf_wei <- pweibull(x_seq, shape = x_wei$estimate[1],
                    scale = x_wei$estimate[2])
    cdf_df <- cdf_df %>% mutate(cdf_weibull = x_cdf_wei)
  }
  if ("logistic" %in% distributions) {
    x_logis <- fitdistr(x, "logistic")
    x_cdf_logis <- plogis(x_seq, x_logis$estimate[1],
                          x_logis$estimate[2])
    cdf_df <- cdf_df %>% mutate(cdf_logistic = x_cdf_logis)
  }

  return(cdf_df)
}

```

## Multi-Distribution Kolmogorov-Smirnov Test
### Description
### Usage
### Arguments
### Details
```{r multiKS}

multiKS <- function(x, distributions) {
# check if "all" was passed to distributions
  if ("all" %in% distributions) {
    distributions <- c("normal",
                       "lognormal",
                       "gamma",
                       "exponential",
                       "cauchy",
                       "t",
                       "weibull",
                       "logistic")
  }
  KS_df <- data.frame(matrix(ncol=3, nrow=0))
  colnames(KS_df) <- c("Distribution", "Distance", "P-Value")
  # check normal
  if ("normal" %in% distributions) {
    x_n <- fitdistr(x, "normal")
    x_KS_n <- ks.test(x, "pnorm", mean=x_n$estimate[1],
                   sd = x_n$estimate[2])
    KS_n <- data.frame(matrix(ncol=0, nrow=1))
    KS_n$Distribution <- "Normal"
    KS_n$Distance <- if (is.null(x_KS_n$statistic)
                         == FALSE) {x_KS_n$statistic}
                     else {"NA"}
    KS_n$PValue <- if (is.null(x_KS_n$p.value)
                       == FALSE) {x_KS_n$p.value}
                   else {"NA"}
    KS_df <- rbind(KS_df, KS_n)
  }
  if ("lognormal" %in% distributions) {
      x_ln <- fitdistr(x, "lognormal")
      x_KS_ln <- ks.test(x, "plnorm",
                         meanlog=x_ln$estimate[1],
                   sdlog = x_ln$estimate[2])[c(1, 2)]
    KS_ln <- data.frame(matrix(ncol=0, nrow=1))
    KS_ln$Distribution <- "Lognormal"
    KS_ln$Distance <- if (is.null(x_KS_ln$statistic)
                         == FALSE) {x_KS_ln$statistic}
                     else {"NA"}
    KS_ln$PValue <- if (is.null(x_KS_ln$p.value)
                       == FALSE) {x_KS_ln$p.value}
                   else {"NA"}
    KS_df <- rbind(KS_df, KS_ln)
  }
  if ("gamma" %in% distributions) {
    x_g <- fitdistr(x, "gamma")
    x_KS_g <- ks.test(x, "pgamma", shape=x_g$estimate[1],
                    rate=x_g$estimate[2])
    KS_g <- data.frame(matrix(ncol=0, nrow=1))
    KS_g$Distribution <- "Gamma"
    KS_g$Distance <- if (is.null(x_KS_g$statistic)
                         == FALSE) {x_KS_g$statistic}
                     else {"NA"}
    KS_g$PValue <- if (is.null(x_KS_g$p.value)
                       == FALSE) {x_KS_g$p.value}
                   else {"NA"}
    KS_df <- rbind(KS_df, KS_g)
  }
  if ("exponential" %in% distributions) {
    x_exp <- fitdistr(x, "exponential")
    x_KS_exp <- ks.test(x, "pexp", rate = x_exp$estimate)
    KS_exp <- data.frame(matrix(ncol=0, nrow=1))
    KS_exp$Distribution <- "Exponential"
    KS_exp$Distance <- if (is.null(x_KS_exp$statistic)
                         == FALSE) {x_KS_exp$statistic}
                     else {"NA"}
    KS_exp$PValue <- if (is.null(x_KS_exp$p.value)
                       == FALSE) {x_KS_exp$p.value}
                   else {"NA"}
    KS_df <- rbind(KS_df, KS_exp)
  }
  if ("cauchy" %in% distributions) {
    x_cau <- fitdistr(x, "cauchy")
    x_KS_cau <- ks.test(x, "pcauchy",
                         location = x_cau$estimate[1],
                         scale = x_cau$estimate[2])
    KS_cau <- data.frame(matrix(ncol=0, nrow=1))
    KS_cau$Distribution <- "Cauchy"
    KS_cau$Distance <- if (is.null(x_KS_cau$statistic)
                         == FALSE) {x_KS_cau$statistic}
                     else {"NA"}
    KS_cau$PValue <- if (is.null(x_KS_cau$p.value)
                       == FALSE) {x_KS_cau$p.value}
                   else {"NA"}
    KS_df <- rbind(KS_df, KS_cau)
  }
  if ("t" %in% distributions) {
    x_t <- fitdistr(x, "t")
    x_KS_t <- ks.test(x, "pt", df = x_t$estimate[3])
    KS_t <- data.frame(matrix(ncol=0, nrow=1))
    KS_t$Distribution <- "t"
    KS_t$Distance <- if (is.null(x_KS_t$statistic)
                         == FALSE) {x_KS_t$statistic}
                     else {"NA"}
    KS_t$PValue <- if (is.null(x_KS_t$p.value)
                       == FALSE) {x_KS_t$p.value}
                   else {"NA"}
    KS_df <- rbind(KS_df, KS_t)
  }
  if ("weibull" %in% distributions) {
    x_wei <- fitdistr(x, "weibull")
    x_KS_wei <- ks.test(x, "pweibull",
                        shape = x_wei$estimate[1],
                    scale = x_wei$estimate[2])
    KS_wei <- data.frame(matrix(ncol=0, nrow=1))
    KS_wei$Distribution <- "Weibull"
    KS_wei$Distance <- if (is.null(x_KS_wei$statistic)
                         == FALSE) {x_KS_wei$statistic}
                     else {"NA"}
    KS_wei$PValue <- if (is.null(x_KS_wei$p.value)
                       == FALSE) {x_KS_wei$p.value}
                   else {"NA"}
    KS_df <- rbind(KS_df, KS_wei)
  }
  if ("logistic" %in% distributions) {
    x_logis <- fitdistr(x, "logistic")
    x_KS_logis <- ks.test(x, "plogis", x_logis$estimate[1],
                          x_logis$estimate[2])
    KS_logis <- data.frame(matrix(ncol=0, nrow=1))
    KS_logis$Distribution <- "Loglogistic"
    KS_logis$Distance <- if (is.null(x_KS_logis$statistic)
                         == FALSE) {x_KS_logis$statistic}
                     else {"NA"}
    KS_logis$PValue <- if (is.null(x_KS_logis$p.value)
                       == FALSE) {x_KS_logis$p.value}
                   else {"NA"}
    KS_df <- rbind(KS_df, KS_logis)
  }
  
  KS_df <- KS_df %>%
      mutate(
        Distribution = as.factor(Distribution),
        Distance = as.numeric(Distance),
        PValue = as.numeric(format(as.numeric(PValue),
                        scientific = FALSE))
      )
  KS_df$Distance <- round(KS_df$Distance, 3)
  KS_df$PValue <- round(KS_df$PValue, 3)

  return(KS_df)
}

```


## ggCDF
### Description
Function that generates a CDF plot using ggplot2
### Usage
### Arguments
### Details

```{r ggcdf}

## Example plot
ggplot(cont_cdf(dat$Fv.Fm, 100, "all")) +
    geom_path(aes(x=x_seq, y=dens,
                  color = "Actual Density"), size=3)+
    geom_path(aes(x=x_seq, y=cdf_normal,
                  color="Normal"), size=2)+
    geom_path(aes(x=x_seq, y=cdf_lognormal,
                  color="Lognormal"),size=2)+
    geom_path(aes(x=x_seq, y=cdf_gamma,
                  color="Gamma"), size=2)+
    geom_path(aes(x=x_seq, y=cdf_exponential,
                  color="Exponential"),size=2)+
    geom_path(aes(x=x_seq, y=cdf_cauchy,
                  color="Cauchy"), size=2)+
    geom_path(aes(x=x_seq, y=cdf_t,
                  color="t"),size=2)+
    geom_path(aes(x=x_seq, y=cdf_weibull,
                  color="Weibull"), size=2)+
    geom_path(aes(x=x_seq, y=cdf_logistic,
                  color="Loglogistic"), size=2)+
    theme_bw()+
    scale_color_scico_d(begin=0, end=0.8,
                        palette = a_palette)+
    scale_shape_discrete()+
    guides(color = guide_legend(title="Distribution"))+
  labs(title = "Fv/Fm Cumulative Distribution Function
       (PDF) Plot")+
    ylab("CDF")+
    xlab("Fv/Fm")+
    theme(
      text = element_text(size=24, family="mont"),
      legend.position="right",
      legend.position.inside = c(.85,.6),
      axis.title = element_text(size=24,
                    family = "mont", face= "bold"),
      title = element_text(size=24,
                           family = "open", face= "bold")
    )

## Plot function

```
